\documentclass[12pt]{article}

\newcommand{\collaborators}{Aksel Kretsinger-Walters (\texttt{adk2164}), Alena Chan
(\texttt{ac5477}), Andrey Aksyutkin (\texttt{aa5499}), Blake Sisson (\texttt{mbs2246})}

\makeatletter
\def\input@path{{../}{../../}{../../../}} % add as many parents as you need
\makeatother
\input{pset_template.tex} %% DO NOT CHANGE THIS LINE

% Override the template margins to be less than 1 inch
% Load geometry after the template so it takes precedence.
\usepackage[margin=0.6in]{geometry}

\title{Waymo Fleet Profitability Optimizer}
\author{
		Aksel Kretsinger-Walters \and
		Alena Chan \and
		Andrey Aksyutkin \and
		Blake Sisson \footnote{Collaborator(s): \collaborators}
}
\date{}
\begin{document}
\maketitle

\subsection*{Problem Statement}
One of the most promising and revolutionary applications of reinforcement learning is in
the domain of autonomous robots, specifically self-driving cars logistics. The challenges facing the
autonomous car community span multiple dimensions: including intellectual, ethical, financial,
and technical hurdles. For our project, we narrow our focus on optimizing the profitability
of robo-taxi fleets.

Monitoring, maintaining, and optimizing a large fleet of self driving cars is a complex
problem, and one can quickly think of many dimensions that the problem takes on.
Predicting demand, scheduling maintenance, recharging vehicles, setting competitive
prices, maximizing coverage, minimizing wait times, and more are all separately non-trivial
problems. Jointly optimizing across all of these dimensions and adapting to distribution shifts
is an even more challenging problem, and the interactive nature of the problem makes it a natural
fit for reinforcement learning and agentic approaches.

\subsection*{RL formulation}

We plan to simulate a fleet of self-driving cars as a Markov Decision Process (MDP), and
develop reinforcement learning algorithms to optimize the fleet's operations and profitability.

The state includes the number of cars available in each city region, the number of
current rides, weather conditions, high-low intraday demand conditions, and fuel reserve
for each car, distance to the nearest charge station. Action includes where to send the
car when it is not with a user inside, when to go to gas station, and what price to
demand for each user/district to balance demand.

\subsection*{Interest and Relevance}
This is an area of active research and development, with many companies investing heavily in
self-driving technology. For this technology to reach the market, it is critical to resolve the
safety and reliability challenges that currently exist; however, the economic viability must
also be solved for self-driving to reach its full potential. Given rapid introduction of
Waymo in multiple cities, this problem is increasingly relevant.

For the purposes of this course, we believe that the interactive nature of the
ride-sharing environment,
the high dimensionality of the state and action spaces, the data and research publicly
available, and the potential of performing far better than a heuristic algorithm make the
robotaxi optimization challenge a good fit for our semester project.

\subsection*{Previous Research}
Ride-sharing in general is an active area of research in the reinforcement learning
community \cite{ridesharing_survey}. There are existing data sets and open simulation environments.

\subsection*{Environment Data}

Our research is based on publicly available datasets to model crowd behavior, weather
conditions, and seasonal effects.

To model crowd behavior, we will utilize data from the Longitudinal Employer-Household
Dynamics (LEHD) program of the U.S. Census Bureau. The LEHD data provide valuable
information on workforce dynamics, which can serve as a proxy for crowd movement and
density. Specifically, we will leverage the following public-use data products:
\begin{itemize}
		\item \textbf{LEHD Origin-Destination Employment Statistics (LODES):} This data set
				provides detailed spatial information on job locations and worker residential
				locations, which is crucial to understanding commuting patterns and daytime
				population distributions.
		\item \textbf{Quarterly Workforce Indicators (QWI):} QWIs offer statistics on
				employment, job creation, and earnings for various demographic groups and industries.
				This data will be used to analyze the composition and temporal dynamics of the workforce.
\end{itemize}

For weather conditions, we will use historical meteorological data from the National
Weather Service. This data will include variables such as temperature, precipitation, and
wind speed.

\subsection*{Proposed solution}
We use a simulation environment to produce states (traffic, demand, weather conditions)
and develop reward model to balance wait times, profit and coverage. We plan to compare
multiple algorithms: policy gradient, mulit-agent RL and other. We will evaluate
efficiency of our approach against heuristic distribution policies.

\begin{thebibliography}{9}

		\bibitem{lehd} U.S. Census Bureau. "LEHD Origin-Destination Employment Statistics
		(LODES)." \url{https://lehd.ces.census.gov/data/}
		\bibitem{suttonbarto} Sutton, R. S., \& Barto, A. G. "Reinforcement Learning: An
		Introduction." MIT Press, 2018.
		\bibitem{waymoopen} Waymo. "Waymo Open Dataset." \url{https://waymo.com/open/}
		\bibitem{citibike_rl} Xiao, I. "Reinforcement Learning Project: CitiBike."
		\url{https://github.com/ianxxiao/reinforcement_learning_project/blob/master/Reports/Presentation_RL_citiBike_20180514.pdf}
		\bibitem{ridesharing_survey} Li, J., Li, X., \& Wang, F. "Reinforcement Learning for
		Ridesharing: An Extended Survey." arXiv preprint arXiv:2102.11896, 2021.
		\bibitem{lin2018} Lin, K., Zhou, M., Wang, J., \& Li, L. (2018). \textit{Efficient
		Large-Scale Fleet Management via Multi-Agent Deep Reinforcement Learning}. Proceedings
		of the 24th ACM SIGKDD International Conference on Knowledge Discovery \& Data Mining (KDD 2018).
		\bibitem{kim2020} Kim, H., \& Kim, Y. (2020). \textit{Optimizing Large-Scale Fleet
				Management on a Road Network using Multi-Agent Deep Reinforcement Learning with Graph
		Neural Network}. arXiv preprint arXiv:2011.06175. \url{https://arxiv.org/abs/2011.06175}
		\bibitem{omahony2015} O’Mahony, E., \& Shmoys, D. (2015). \textit{Data Analysis and
		Optimization for (Citi)Bike Sharing}. Proceedings of the Twenty-Ninth AAAI Conference
		on Artificial Intelligence (AAAI 2015).
		\bibitem{xu2021} Xu, Y., Zhang, R., Li, X., \& Wang, F. (2021). \textit{Towards More
		Efficient Shared Autonomous Mobility: A Learning-Based Fleet Repositioning Approach}.
		Proceedings of the AAAI Conference on Artificial Intelligence, 35(1), 1166-1174.
		\url{https://ojs.aaai.org/index.php/AAAI/article/view/16033}

		% https://www.sciencedirect.com/science/article/pii/S2772424724000246?via%3Dihub
\end{thebibliography}
\end{document}
